%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                      %
%     File: Thesis_Acknowledgments.tex                                 %
%     Tex Master: Thesis.tex                                           %
%                                                                      %
%     Author: Andre C. Marta                                           %
%     Last modified :  2 Jul 2015                                      %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{\acknowledgments}\label{sec:acknowledgments}

% Add entry in the table of contents as section
\addcontentsline{toc}{section}{\acknowledgments}

% I might want to re-read this later and make some changes (to the content, not the formatting).

First and foremost, I would like to express my deepest gratitude to my research supervisors, Prof. Yasser Omar and Prof. Zoltán Zimborás. Their invaluable guidance and support have been instrumental in shaping my research skills and scientific thinking. I am profoundly grateful for the opportunity to work with them and for the trust they have placed in me. Additionally, I am especially thankful to Prof. Yasser Omar for providing me with opportunities to broaden my horizons through international collaborations and participation in various European projects.

% Removed: ", if not more so,".
Equally important are the contributions of my colleagues Bence Bakó and Miguel Murça. Their willingness to discuss ideas and provide feedback whenever I felt lost was a tremendous support. Without their assistance, this thesis would not be what it is today. To them, I extend my heartfelt thanks.

I would also like to extend my gratitude to my family, who have always been there for me, offering advice on both thesis-related matters and beyond.  Special thanks to my sister, Carolina, who has listened to me rant for hours on end about random quantum computing topics that she has absolutely no clue about. Although I suspect (know) that sometimes she was just pretending to listen, I appreciate the effort nonetheless.

Lastly, I would like to thank my closest friends for their unwavering support and encouragement. The small discussions with you often led to new ideas and insights, helping me view my work from different perspectives. Besides, having someone to play padel with and relax was always a welcome respite. Thank you for your companionship.

\vspace*{\fill}

This work was developed in collaboration with Zoltán Zimborás and Bence Bakó, from the Wigner Research Centre for Physics (Hungary), in the context of project: \textit{HQCC – Hybrid Quantum-Classical Computing}, supported by the EU QuantERA ERA-NET Co-fund in Quantum Technologies and by FCT -- Funda\c{c}\~{a}o para a Ci\^{e}ncia e a Tecnologia (QuantERA/004/2021).

% I should ask Yasser where to put this 'Acknowledgement'. I think it's fine here, but I'm not sure.